---
title: "COMP1013_22176385_Assigment"
output: word_document
date: "2025-05-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Identity

## Student Surname: Prema Santhi

## Student Firstname: I Gusti Ayu Agung

## Student ID: 22176385

## Unit Name: Analytics Programming

## Unit Number: COMP1013

# Dataset information
In this assigment, we use several provided datasets such as below:

1. customers.csv

2. geolocation.csv

3. order_items.csv

4. orders.csv

5. payments.csv

6. products.csv

7. sellers.csv

## Load all the dataset
Before begin to the analysis, load all the data first
```{r}
customer.data=read.csv("customers.csv")
geo.data=read.csv("geolocation.csv")
item.data=read.csv("order_items.csv")
orders=read.csv("orders.csv")
payment.data=read.csv("payments.csv")
product.data=read.csv("products.csv")
sellers.data=read.csv("sellers.csv")
```

## Call the necessery libraries
```{r, results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
```
